\chapter{Introduction}
\begingroup
\justifying
\setlength{\parindent}{0pt}
\setstretch{1.5}
\setlength{\parskip}{0.5\baselineskip}
\titlespacing{\chapter}{0pt}{0pt}{0pt}
\titlespacing{\section}{0pt}{0pt}{0pt}

Electronic voting systems offer the prospect of rapid tabulation, broader voter accessibility, and auditable records that can withstand independent scrutiny. These advantages have sustained interest in electoral technology reform for several decades.

Realising those benefits in practice, however, has proved considerably more difficult than anticipated. A viable electronic voting system must simultaneously preserve ballot secrecy, enforce correct state transitions, afford sufficient transparency for independent verification, and remain operationally practical for both election officials and voters. Satisfying all of these requirements at once introduces fundamental design tensions, which account for the persistent shortcomings observed in deployed systems.

Centralised platforms address usability and throughput effectively, yet they concentrate the entire basis of trust in backend operators. Those operators alone govern the audit logs, manage state transitions, and produce the final tallies. Independent verification extends only as far as the operators permit, leaving external auditors with no publicly verifiable recourse in disputed elections.

This dissertation explores an alternative approach in which the basis of assurance shifts from institutional trust to on-chain enforcement underpinned by cryptographic proofs. The architecture integrates three complementary mechanisms within a single end-to-end pipeline: ElGamal homomorphic encryption on the BabyJubJub curve for privacy-preserving tally computation, Groth16 zero-knowledge proofs for succinct on-chain verification of ballot validity and decryption correctness, and commitment-based Merkle inclusion proofs that enable post-election ballot-inclusion verification. Key custody remains the sole explicit trust assumption; all other assurance properties derive from the cryptographic protocol and its on-chain enforcement. The residual trust boundary and its practical implications are examined in detail in later chapters.

The work extends beyond a theoretical design. The resulting prototype compiles, deploys on Ethereum-compatible test networks, and has been validated through a suite of automated tests covering both cryptographic correctness and contract-level behaviour.

\section{Problem Statement}

Designing a voting system that minimises reliance on trusted operators entails reconciling a set of requirements that are not naturally aligned.

\textit{Ballot confidentiality} must be maintained throughout submission, on-chain storage, and tallying. \textit{Correctness verification} must confirm both the well-formedness of individual ballots and the integrity of the aggregate tally, without disclosing the content of any ballot. \textit{State-transition integrity} demands tamper-evident, on-chain enforcement of the election lifecycle. Finally, \textit{voter-side auditability} requires a verification path that enables each voter to confirm the inclusion of their ballot using only a locally stored receipt and a published audit bundle, with the bundle root checked for consistency against the on-chain \texttt{merkleRoot} and with no address-based lookup during verification.

Let $V$ denote the number of cast ballots and $N$ the number of candidates. Every accepted ballot must encode precisely one valid candidate choice. The homomorphic aggregate of all encrypted ballots must yield a consistent tally, and the published result must be accompanied by cryptographic evidence that any independent observer can verify.

\section{Motivation}

Three convergent technical advances make the proposed combination of methods practically feasible.

First, blockchain platforms such as Ethereum provide programmable smart contracts together with append-only event logs and deterministic execution semantics. These properties eliminate an entire class of attacks common to centralised systems---namely, silent, undetectable modifications to election records \cite{nakamoto2008,wood2014}.

Second, recent progress in zk-SNARK constructions, most notably the Groth16 proving system, enables the generation of succinct, constant-size proofs whose on-chain verification cost remains bounded within practical smart-contract gas limits \cite{groth2016}.

Third, algebraically efficient hash functions such as Poseidon, designed specifically for arithmetic circuits, together with Merkle-tree constructions, support efficient post-election inclusion checks over large ballot sets without imposing prohibitive on-chain storage costs \cite{poseidon2021,merkle1988}.

Taken together, these components open a viable design space in which the tension between voter privacy and public verifiability becomes substantially more manageable. Assurance can be grounded in cryptographic proofs and publicly auditable execution traces rather than in institutional trust alone---though, as the evaluation in Chapter~5 will show, closing every residual gap requires further engineering effort beyond the scope of this prototype.

\section{Research Objectives}

This study pursues five objectives.

\begin{itemize}
    \item Design and implement a contract-level state machine that enforces explicit, tamper-evident lifecycle transitions for the election.
    \item Enable private ballot submission through one-hot ElGamal encryption on the BabyJubJub curve, ensuring that each encrypted ballot encodes exactly one valid candidate choice.
    \item Employ Groth16 zero-knowledge proofs for on-chain verification of both individual ballot validity and the correctness of tally claims derived from off-chain homomorphic aggregation.
    \item Develop a receipt-based audit mechanism whereby voters can confirm, using a locally stored receipt and a published audit bundle, that their ballot commitment has been correctly included. The verification procedure checks the consistency between the bundle root and the on-chain \texttt{merkleRoot}, and requires no address-based lookup during the verification step.
    \item Evaluate the strengths of the resulting implementation alongside its limitations, including residual engineering constraints, trust assumptions, and binding gaps that currently prevent higher-assurance deployment.
\end{itemize}
\section{Research Questions}

Three questions guide this study: feasibility, comparative evaluation, and gap analysis.

\textbf{RQ1:} \textit{Is it feasible to integrate encrypted one-hot ballots and zero-knowledge-proof-based checks into a browser-to-blockchain workflow without relying on trusted middleware?}

Can encryption, proof generation, and smart-contract verification operate as a single browser-launched pipeline, with no middleware positioned to tamper with ballots or interfere with verification? Chapter 4 details the implementation, and Chapter 5 evaluates it through scenario-based testing.

\textbf{RQ2:} \textit{To what extent does the proposed pipeline improve verifiability over centralized log-based auditing, particularly by enabling independent tally replay and ballot-commitment inclusion verification without server interaction?}

Two tests follow: can an independent party recompute the tally from on-chain ciphertexts and reach the same result, and can a voter confirm inclusion without contacting any server? Chapter 2 frames the comparison, and Chapter 5 provides the security analysis.

\textbf{RQ3:} \textit{Which intended security guarantees are fully achieved in the current implementation, and which remain incomplete due to practical engineering constraints?}

RQ3 separates what the prototype enforces---such as on-chain format checks rejecting malformed ballots---from what it only partially addresses, such as binding integrity between voter intent and the committed payload. Evidence appears in Chapter 5 and the traceability appendix.

Table~\ref{tab:rq-mapping} indicates where each question is primarily addressed.

\begin{table}[htbp]
\centering
\small
\caption{Research Question to Validation Mapping}
\label{tab:rq-mapping}
\begin{tabular}{L{2.0cm} L{5.0cm} L{5.8cm}}
\toprule
RQ & Evaluation Focus & Primary Evidence \\
\midrule
RQ1 & End-to-end integration feasibility & Chapter 4 (implementation) and Chapter 5 (scenario validation) \\
RQ2 & Comparative verifiability via tally replay and server-free ballot-commitment verification & Chapter 2 (positioning) and Chapter 5 (security analysis) \\
RQ3 & Gap analysis of achieved and outstanding security properties & Chapter 5 (outstanding issues) and the traceability appendix \\
\bottomrule
\end{tabular}
\end{table}
\section{Scope and Boundaries}

This e-voting system is a reproducible research prototype, targeting deployment of a two-candidate election.

Circuits are compiled and tested for a fixed candidate count of $N = 2$; larger $N$ requires significantly more engineering effort and is not part of this prototype. Voter eligibility relies on an administrator-managed whitelist whose real-world correctness is outside the protocol's cryptographic scope.

\textbf{Trust assumptions.} The prototype assumes a semi-trusted administrator who follows the protocol. The admin keeps the private key safely, gives an honest whitelist, and would not release wrong results since it can be easily identified by the public. For on-chain results, people only need to trust cryptography and no institutional trust is needed.

Out of scope: coercion resistance; threshold key management; decentralized governance; formal contract verification; cross-election voter privacy.

\section{Contributions}

This work produces four concrete contributions:

\begin{itemize}
    \item A cross-stack prototype spanning Solidity smart contracts, Circom zero-knowledge circuits, browser-side JavaScript for cryptographic computation and wallet interaction, and Python for local serving and cryptographic validation. The prototype compiles, deploys to Ethereum-compatible networks, and is accompanied by 76 automated tests covering both contract behaviour and underlying cryptographic primitives.
    \item A vote-submission mechanism that cryptographically binds each ballot's Poseidon commitment to the corresponding encrypted payload through a shared ciphertext-binding hash, both of which serve as public inputs to the vote proof. This binding reduces the risk of discrepancies between the statement proved by the zero-knowledge proof and the data recorded on-chain.
    \item A tally pipeline that performs homomorphic ciphertext aggregation off-chain, verifies decryption correctness through an on-chain Groth16 proof, and anchors the finalised result set with a published Merkle root of all ballot commitments.
    \item A receipt-based audit path that enables each voter to confirm ballot inclusion using a locally stored receipt and an exported audit bundle, with no address-based lookup required during the verification step. The verification procedure checks commitment membership via a Merkle inclusion proof and validates root consistency against the on-chain \texttt{merkleRoot}.
\end{itemize}

\section{Dissertation Structure}

The remainder of this dissertation is organised as follows.

Chapter~2 surveys related work in electronic voting and establishes the cryptographic foundations---homomorphic encryption, zero-knowledge proofs, and Merkle-based auditability---on which the protocol relies.

Chapter~3 presents the system model, defines the protocol roles and threat model, specifies the election-lifecycle state machine, and details the ballot encoding, proof constraints, and audit construction.

Chapter~4 describes the implementation, covering smart-contract design, circuit architecture, frontend engineering, testing strategy, and deployment procedures.

Chapter~5 evaluates the prototype through automated test results, on-chain gas measurements, proof-generation latency, requirement traceability, security analysis, and an assessment of outstanding engineering gaps.

Chapter~6 summarises the findings, revisits the research questions, and proposes a phased roadmap for future hardening.

Two appendices supplement the main text: Appendix~A provides a step-by-step guide for reproducing the prototype locally, and Appendix~B presents the requirement traceability checklist with acceptance status.

\endgroup
