\chapter{Results, Auditability, and Security Analysis}
\begingroup
\justifying
\setlength{\parindent}{0pt}
\setstretch{1.5}
\setlength{\parskip}{0.5\baselineskip}
\titlespacing{\chapter}{0pt}{0pt}{0pt}
\titlespacing{\section}{0pt}{0pt}{0pt}

\section{Evaluation Methodology}

Evaluation focuses on three aspects:
\begin{itemize}
    \item \textbf{Functional correctness}: whether the system supports the intended election workflow end to end.
    \item \textbf{Verifiability}: whether invalid ballots and unsupported tally claims are rejected by contract verification.
    \item \textbf{Auditability and privacy}: whether voters can verify inclusion without revealing identity or ballot content.
\end{itemize}

Two complementary strategies are used: automated testing (contract and cryptographic unit tests) and workflow-based validation using the administrator and voter clients.

\section{Automated Verification Results}

\subsection{Smart Contract Integration Tests}

Hardhat tests validate:
\begin{itemize}
    \item voter registry permissions and one-vote enforcement;
    \item election lifecycle restrictions (Created/Active/Ended/Tallied);
    \item vote casting success path with Groth16 proof verification;
    \item invalid voting attempts (unregistered voter, duplicate vote, wrong phase);
    \item tally submission and state finalization behavior.
\end{itemize}

\subsection{Cryptographic Unit Tests}

Python tests validate BabyJubJub ElGamal primitives:
\begin{itemize}
    \item group arithmetic (associativity, identity, negation);
    \item key generation and serialization;
    \item encryption/decryption correctness across small message range;
    \item homomorphic addition correctness;
    \item one-hot encoding correctness for fixed candidate cardinality;
    \item bounded discrete-log recovery for tally extraction.
\end{itemize}

At the time of evaluation, the local test baseline was:
\begin{itemize}
    \item Hardhat tests: 33 passed.
    \item Python crypto tests: 43 passed.
\end{itemize}

The outcomes provide evidence to conclude that the prototype is replicable, and functional at a high level of coherence under controlled settings.

\subsection{Runtime and Reproducibility Snapshot}

To improve reproducibility, this dissertation records a concrete execution snapshot from the current repository revision.

\begin{table}[htbp]
\centering
\small
\caption{Automated Test Runtime Snapshot (Local Evaluation Run)}
\label{tab:runtime-snapshot}
\begin{tabular}{L{2.8cm} L{1.8cm} L{2.0cm} L{5.5cm}}
\toprule
Suite & Passed & Runtime & Command \\ 
\midrule
Hardhat contract/integration tests & 33/33 & 20 s & \texttt{npm test} \\
Python cryptography unit tests & 43/43 & 27.59 s & \texttt{python -m pytest tests/ -q} \\ 
\bottomrule
\end{tabular}
\end{table}

Both suites completed with zero failures, for a combined 76 passing tests. This does not prove security in adversarial deployment, but it gives strong evidence that the implementation claims in Chapters 3--4 are executable and internally consistent on a fresh local run.

Table~\ref{tab:evaluation-scenarios} provides a scenario-oriented view of what was validated and what evidence was observed.

\begin{table}[htbp]
\centering
\small
\caption{Evaluation Scenarios and Outcomes}
\label{tab:evaluation-scenarios}
\begin{tabular}{L{4.0cm} L{2.6cm} L{6.6cm}}
\toprule
Scenario & Result & Evidence \\
\midrule
Registered voter submits valid proof & Pass & Vote accepted and \texttt{EncryptedVoteCast} emitted. \\
Unregistered voter attempts to vote & Pass (rejected) & Contract revert on registration check before state change. \\
Duplicate vote attempt & Pass (rejected) & Contract revert through one-vote gating logic. \\
Tally finalization with proof & Pass & Result submission accepted after verifier call and status transition. \\
Receipt-based inclusion verification & Pass & Voter-side proof validation succeeds with matching receipt and bundle entry. \\
\bottomrule
\end{tabular}
\end{table}

\section{End-to-End Workflow Validation}

A practical end-to-end run proceeds:
\begin{enumerate}
    \item Administrator adds candidates and registers voter addresses.
    \item Election is started; voters generate ciphertexts and vote proofs in browser.
    \item Contract verifies vote proofs and emits encrypted vote events.
    \item Administrator aggregates encrypted ballots, decrypts totals, generates tally proof.
    \item Contract verifies tally proof and finalizes election with published Merkle root.
    \item Administrator exports \texttt{audit\_bundle.json}; voters verify inclusion using receipt.
\end{enumerate}

The main working result is that, for a voter to verify inclusion without invoking \texttt{getVoteRecord(address)} thereby reducing address based linkability which satisfies the project requirement of provable privacy.

\section{Performance Evaluation}

This section reports some quantitative measurements about the system's on-chain gas cost and ZK proof circuit size, client-side artifact footprint, and smart contract bytecode size. All measurements were taken on the project's current build artifacts with Solidity 0.8.20 (optimizer enabled, 200 runs), Groth16 on BN254, and the circuit configuration of $N=2$ candidates.

\subsection{Circuit Complexity}

Table~\ref{tab:circuit-complexity} summarizes the R1CS constraint systems produced by a fresh local Circom recompilation of the current repository revision. These values were extracted from the generated \texttt{.r1cs} files using \texttt{snarkjs r1cs info}.

\begin{table}[htbp]
\centering
\small
\caption{ZKP Circuit Complexity (N = 2 Candidates)}
\label{tab:circuit-complexity}
\begin{tabular}{L{3.2cm} L{2.0cm} L{2.0cm} L{2.0cm} L{2.0cm}}
\toprule
Circuit & Constraints & Wires & Public Inputs & Private Inputs \\
\midrule
\texttt{vote\_proof} & 14{,}590 & 14{,}593 & 4 & 12 \\
\texttt{tally\_proof} & 16{,}786 & 16{,}790 & 15 & 1 \\
\bottomrule
\end{tabular}
\end{table}

Both circuits operate on the BN254 (also referred to as BN128) curve. The vote proof circuit encodes commitment verification, candidate range checks, one-hot encryption validity for $N$ candidates, and ciphertext hash binding. The tally proof circuit encodes key ownership, per-candidate decryption correctness, and sum consistency. Constraint counts scale linearly with $N$ due to per-candidate ElGamal verification sub-circuits.

\subsection{Client-Side Artifact Footprint}

Proof generation in the browser requires downloading WASM witness generators and proving key files. Table~\ref{tab:artifact-sizes} reports the sizes of artifacts that must be loaded on the client side.

\begin{table}[htbp]
\centering
\small
\caption{ZKP Artifact Sizes}
\label{tab:artifact-sizes}
\begin{tabular}{L{4.5cm} L{2.5cm} L{5.5cm}}
\toprule
Artifact & Size & Role \\
\midrule
\texttt{vote\_proof.wasm} & 2.94 MiB & Witness generator for vote proof \\
\texttt{vote\_proof\_final.zkey} & 7.11 MiB & Proving key for vote proof \\
\texttt{tally\_proof.wasm} & 143.0 KiB & Witness generator for tally proof \\
\texttt{tally\_proof\_final.zkey} & 9.12 MiB & Proving key for tally proof \\
\texttt{vote\_proof.r1cs} & 2.79 MiB & Constraint system (build-time only) \\
\texttt{tally\_proof.r1cs} & 3.24 MiB & Constraint system (build-time only) \\
\bottomrule
\end{tabular}
\end{table}

For a voter casting a ballot, the browser must load approximately 10.06~MiB of proof artifacts (WASM + zkey). One download per session but cacheable; it is around 9.3~MiB for a tallying admin. These numbers line up with typical Groth16 deploys at this size of constraints and are well within reasonable budgetary expectations in terms of browser download sizes for modern links.

\subsection{On-Chain Gas Consumption}

Gas costs are at the core when assessing the viability of a protocol based upon blockchains. Due to the \texttt{ecPairing} precompile, Groth16 proof verification takes up most of the cost on chain, which on EVM costs $45{,}000 + 34{,}000 \times k$ gas for $k$ pairing checks (Groth16 uses $k = 4$, yielding 181{,}000 gas for the pairing alone).

Table~\ref{tab:gas-costs} reports measured gas consumption from transaction receipts in a fresh local Hardhat run against the current repository revision. The dominant cost components are interpreted with reference to EIP-1108 precompile costs \cite{eip1108} and the storage-access behavior of \texttt{Voting.sol}.

\begin{table}[htbp]
\centering
\small
\caption{Measured Gas Consumption Per Operation (Local Hardhat Run)}
\label{tab:gas-costs}
\begin{tabular}{L{4.5cm} L{2.5cm} L{5.5cm}}
\toprule
Operation & Measured Gas & Dominant Cost Component \\
\midrule
\texttt{castVote} (voter) & 367{,}494--401{,}682 & Groth16 verification ($\sim$181k) + vote record writes + encrypted payload event emission \\
\texttt{updateTallyResults} (admin) & 498{,}307 & Groth16 verification ($\sim$181k) + tally-result writes + Merkle-root finalization + events \\
\texttt{addCandidates} & 127{,}649 & Two candidate writes + candidateCount update + events \\
\texttt{startElection} & 31{,}924 & Status transition write \\
\texttt{endElection} & 29{,}839 & Status transition write \\
\texttt{registerVoter} (single) & 75{,}219--92{,}331 & Registration mapping write; the first registration on a fresh deployment is highest \\
\bottomrule
\end{tabular}
\end{table}

To contextualize these costs: at a gas price of 30~gwei and an ETH price of \$2{,}500, one \texttt{castVote} transaction costs approximately 0.01102--0.01205~ETH (\$27.6--\$30.1), and one \texttt{updateTallyResults} costs approximately 0.01495~ETH (\$37.4). The Ethereum block gas limit (30M gas) can accommodate roughly 74--81 vote transactions per block if they were the only transactions, although in reality the blockspace is shared. (Gas price and ETH price numbers are representative prices as evaluated, they will fluctuate based on network performance.)

\subsection{Proof Generation Latency}

Groth16 proof generation is the most computation intensive part on the client side. It includes multi-scalar multiplication over the proving key, which is proportional to the number of circuit constraints. For the current configuration ($N = 2$), a fresh local Node.js measurement on the current repository revision produced vote-proof latencies between 615~ms and 1{,}052~ms across three representative ballots, and a tally-proof latency of about 848~ms. Browser-based execution using snarkjs WASM is expected to be similar or somewhat slower depending on device hardware and caching behavior.

These times are acceptable for a prototype where voters submit one ballot per election. However, they represent a usability consideration: the voter interface should communicate proof-generation progress to avoid the impression of an unresponsive application.

\subsection{Contract Bytecode Size}

Ethereum imposes a 24{,}576-byte limit on deployed contract bytecode (EIP-170 \cite{eip170}). Table~\ref{tab:bytecode-sizes} confirms that all contracts remain well within this limit.

\begin{table}[htbp]
\centering
\small
\caption{Deployed Contract Bytecode Sizes}
\label{tab:bytecode-sizes}
\begin{tabular}{L{3.5cm} L{2.5cm} L{6.5cm}}
\toprule
Contract & Deployed Size & Notes \\
\midrule
\texttt{Voting} & 10{,}168 bytes (9.9 KB) & Main election process \\
\texttt{VoteVerifier} & 1{,}658 bytes (1.6 KB) & Groth16 verifier (snarkjs-generated) \\
\texttt{TallyVerifier} & 2{,}690 bytes (2.6 KB) & Groth16 verifier (snarkjs-generated) \\
\bottomrule
\end{tabular}
\end{table}

The total amount of deployed bytecode across all 3 contracts is 14{,}516 bytes. The Voting contract is large (41\%) but still well under the EIP-170 limit and leaves space for more features if needed. Combining the voter register in with the voting contract results in a slightly larger contract, but removed two deployments. The Groth16 verifier contracts are small since snarkjs generates optimised inline pairing-check code.

\subsection{Scalability Discussion}

The scalability properties of the system are different for each component. The on-chain gas used per vote (\texttt{castVote}) is almost constant, irrespectively from how much voting has already happened, since every ballot is checked separately and recorded. Let $B_{dlog}$ denote the discrete-log search bound used by the tally client. In the current implementation, $B_{dlog}=V+100$ where $V$ is total votes cast. Several operations scale with election size:

\begin{itemize}
    \item \textbf{Tally aggregation} requires $O(VN)$ elliptic-curve point additions, performed off-chain by the administrator. For $V = 1{,}000$ votes and $N = 2$ candidates, this amounts to 2{,}000 point additions, completing in under one second on modern hardware.
    \item \textbf{Discrete-log recovery} during tally decryption uses runtime baby-step giant-step. Per candidate, complexity is $O(\sqrt{B_{dlog}})$; for an election with $N$ candidates, full tally recovery is $O(N\sqrt{B_{dlog}})$. No persistent global precomputed table is required; instead, a temporary baby-step map of size $O(\sqrt{B_{dlog}})$ is built in memory during each recovery call. Since per-candidate tallies are bounded above by $V$, the current bound $B_{dlog}=V+100$ is sufficient for correctness but may become a client-side bottleneck for very large elections.
    \item \textbf{Merkle tree construction} is $O(V)$ with $O(\log V)$ verification per inclusion check. For $V = 10{,}000$ commitments, tree construction completes in milliseconds.
    \item \textbf{Circuit constraint scaling}: constraint growth is approximately linear in $N$ for the per-candidate ElGamal sub-circuits, but exact totals require recompilation for each target $N$. Therefore, values beyond $N=2$ in this dissertation are treated as extrapolative rather than measured. In addition, straightforward scaling in the current vote-circuit layout is bounded by the ciphertext-hash arity (\texttt{Poseidon(4N)}), so larger-$N$ support requires circuit-level hash-layout refactoring before claiming concrete higher-$N$ counts.
\end{itemize}

Table~\ref{tab:scalability-summary} summarizes the scaling profile.

\begin{table}[htbp]
\centering
\small
\caption{Scalability Profile by Component}
\label{tab:scalability-summary}
\begin{tabular}{L{4.2cm} L{2.8cm} L{5.5cm}}
\toprule
Component & Complexity & Practical Bound \\
\midrule
Per-vote on-chain cost & $O(1)$ & Constant $\sim$367k--402k gas per vote \\
Tally aggregation (off-chain) & $O(VN)$ & Practical up to $V = 10^5$ \\
DLog recovery (off-chain) & $O(N\sqrt{B_{dlog}})$ & Current admin client uses $B_{dlog}=V+100$ \\
Merkle tree build (off-chain) & $O(V)$ & Practical up to $V = 10^6$ \\
Inclusion verification & $O(\log V)$ & Negligible for all practical $V$ \\
Proof generation (client) & $O(C_{vote})$ & Scales with vote-circuit constraints, where $C_{vote}\approx aN+b$ for fixed circuit design \\
\bottomrule
\end{tabular}
\end{table}

Overall, the prototype is well-suited for small-to-medium elections (up to several thousand voters) in its current form. The primary scaling bottleneck for larger elections would be proof generation latency and artifact download size as candidate count $N$ increases, rather than on-chain gas costs.

\section{Commitment Merkle Auditability}

Keccak256 algorithm is used in the audit design to compute internal nodes via commitment leaves. As a deterministic hash function, there is no randomness in Keccak256; therefore, given the same event set at the same ordering, independent third-party auditors always reproduce the same root value.

\subsection{Audit Bundle Content}

The exported audit bundle includes:
\begin{itemize}
    \item election identifier and chain metadata;
    \item Merkle root;
    \item commitment entries with indices and sibling proofs;
    \item transaction references for traceability.
\end{itemize}

\subsection{Anonymous Verification}

Using local receipt material, the voter client recomputes the commitment, locates matching entry, and verifies Merkle path. A successful verification implies that the commitment is included in the audited set whose root is published.

\section{Requirement Completion and Traceability}

Table~\ref{tab:reqtrace} maps finer-grained requirements to concrete evidence.

\begin{table}[htbp]
\centering
\small
\caption{Fine-Grained Requirement Traceability}
\label{tab:reqtrace}
\begin{tabular}{L{4.0cm} L{2.6cm} L{6.6cm}}
\toprule
Requirement Item & Status & Evidence and Notes \\
\midrule
On-chain election lifecycle enforcement & Completed & Voting state machine and modifiers in \texttt{contracts/Voting.sol}. \\
Voter eligibility and one-vote rule & Completed & Voter registration and one-vote logic integrated directly into \texttt{contracts/Voting.sol}. \\
Ballot confidentiality (no plaintext on-chain) & Completed & Only ciphertext and commitment published; plaintext never stored. \\
Vote legality proof (range + one-hot) & Completed & \texttt{circuits/vote\_proof.circom} + verifier checks in \texttt{castVote}. \\
Ciphertext binding to vote proof & Completed (partial binding) & Circuit proves \texttt{ciphertextHash}; event payload binding not fully enforced (see Section 5.8). \\
Tally correctness proof (decryption) & Completed & \texttt{circuits/tally\_proof.circom} + verifier checks in \texttt{updateTallyResults}. \\
Merkle inclusion auditability & Completed & \texttt{frontend/lib/audit.js}, export bundle, verify inclusion. \\
Voter anonymity in verification & Completed (operational) & Receipt-based verification avoids address lookup; still relies on client privacy. \\
Production readiness & Not completed & Several hardening tasks remain (Section 5.8). \\
\bottomrule
\end{tabular}
\end{table}

\section{Security Analysis}

Figure~\ref{fig:threat-model} summarizes the threat surface considered in this dissertation and the primary control layers used by the implementation.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
    \node[draw, rounded corners, minimum width=5.2cm, minimum height=1.3cm, fill=gray!15, align=center] (core) at (0,0) {E-Voting Protocol Core\\(contracts + circuits + audit logic)};

    \node[draw, rounded corners, minimum width=3.5cm, minimum height=0.95cm] (t1) at (-5.2,2.2) {Malicious voter};
    \node[draw, rounded corners, minimum width=3.5cm, minimum height=0.95cm] (t2) at (5.2,2.2) {Passive observer};
    \node[draw, rounded corners, minimum width=3.5cm, minimum height=0.95cm] (t3) at (-5.2,-2.2) {Compromised admin key};
    \node[draw, rounded corners, minimum width=3.5cm, minimum height=0.95cm] (t4) at (5.2,-2.2) {Client-side malware};

    \draw[->, thick] (t1) -- (core);
    \draw[->, thick] (t2) -- (core);
    \draw[->, thick] (t3) -- (core);
    \draw[->, thick] (t4) -- (core);

    \node[font=\scriptsize, align=center] at (-2.75,1.25) {ZKP vote constraints\\and eligibility gates};
    \node[font=\scriptsize, align=center] at (2.75,1.25) {Ciphertext + commitment\\hide ballot choice};
    \node[font=\scriptsize, align=center] at (-2.7,-1.25) {Tally proof validation\\but key custody risk remains};
    \node[font=\scriptsize, align=center] at (2.7,-1.25) {Outside strict protocol\\guarantee boundary};
\end{tikzpicture}
\caption{Threat Model and Control Mapping}
\label{fig:threat-model}
\end{figure}

\subsection{Integrity Guarantees Achieved}

The system enforces:
\begin{itemize}
    \item lifecycle integrity through explicit state transitions;
    \item ballot validity through on-chain proof verification;
    \item tally proof verification before result finalization.
\end{itemize}

These properties meaningfully reduce risks common to centralized systems where validity and tally logic are opaque.

\subsection{Privacy Guarantees Achieved}

Ballot confidentiality is achieved by encryption and by ensuring that plaintext choices are never submitted to the chain. The commitment receipt is locally stored and does not directly reveal vote choice unless the salt is disclosed.

\subsection{Residual Risks}

The most important residual risks come from key custody and the client's environment. Even if cryptographically correct, an attacked endpoint may leak salt, candidate selection, or manipulate the local witness generation. All this is beyond the guarantee of the protocol.

\section{Outstanding Engineering Gaps and Their Impact}

Code review identified three remaining gaps that affect full assurance. Gaps are classified by impact: \textbf{P1} (high impact, fix before any public deployment) and \textbf{P2} (medium impact, fix before production use):
\begin{itemize}
    \item \textbf{[P1] Result binding gap}: contract verifies tally proof, but does not strictly binds submitted integer result vector to all public signals used by the circuit, which may allows inconsistency as long as no outside check is imposed.
    \item \textbf{[P1] Ciphertext event binding gap}: the emitted encrypted payload is not fully bound to on-chain stored ciphertext hash. If an attacker can influence event payload parsing or indexing, off-chain tally can be poisoned.
    \item \textbf{[P2] Merkle root mutability}: an admin-only root-update function remains available before tally finalization; although overwrites are rejected once the election reaches \texttt{Tallied}, the pre-finalization update path still increases governance attack surface.
\end{itemize}

These issues do not invalidate the prototype's educational value or end-to-end demonstration, but they prevent claiming full compliance with ``fully functional'' in a high-assurance sense.

Figure~\ref{fig:assurance-gap} illustrates the current assurance profile: strong functional completeness, but a smaller unresolved hardening segment that must be closed before production use.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
    \draw[fill=gray!15] (0,0) rectangle (10,1.0);
    \draw[fill=gray!40] (0,0) rectangle (7.6,1.0);
    \draw[thick] (7.6,0) -- (7.6,1.0);
    \node[font=\small] at (3.8,0.5) {Implemented and validated};
    \node[font=\small] at (8.8,0.5) {Hardening gap};

    \node[anchor=west, font=\small] at (0,-0.5) {Illustrative only: core workflow implemented; hardening remains};
\end{tikzpicture}
\caption{Current Assurance Profile (Illustrative)}
\label{fig:assurance-gap}
\end{figure}

\section{Threats to Validity}

The reported findings are subject to standard empirical-study validity limits:
\begin{itemize}
    \item \textbf{Internal validity}: tests are deterministic and mostly run on a local development chain; some network and wallet timing behaviors may be underrepresented.
    \item \textbf{Construct validity}: ``privacy'' in this dissertation primarily means ballot confidentiality plus receipt-based anonymous inclusion verification, not full coercion resistance.
    \item \textbf{External validity}: performance and usability observations may shift under larger electorates, different gas regimes, and different proof-artifact configurations.
    \item \textbf{Conclusion validity}: passing tests indicate consistency with encoded properties, but cannot by themselves establish absence of all protocol or implementation vulnerabilities.
\end{itemize}

To mitigate these threats, the dissertation combines protocol-level reasoning, cross-layer testing, and explicit gap documentation rather than relying on a single evidence source.

\section{Chapter Summary}

This chapter demonstrated that the implemented system is functionally complete as a verifiable privacy-preserving prototype and reproducible under automated testing. Performance evaluation confirmed that on-chain costs are dominated by Groth16 verification and that the system scales adequately for small-to-medium elections. It also provided a requirement-grounded gap analysis that distinguishes current achievement from production-grade assurance.

\endgroup

