\chapter{Implementation}
\begingroup
\justifying
\setlength{\parindent}{0pt}
\setstretch{1.5}
\setlength{\parskip}{0.5\baselineskip}
\titlespacing{\chapter}{0pt}{0pt}{0pt}
\titlespacing{\section}{0pt}{0pt}{0pt}

\section{Project Architecture}

The repository is organised into five major modules:
\begin{itemize}
    \item \texttt{contracts/}: Solidity contracts for election control (including integrated voter registration) and verifier adapters.
    \item \texttt{circuits/}: Circom circuits for vote and tally proofs.
    \item \texttt{frontend/}: voter and administrator web clients, including browser-side cryptographic helpers.
    \item \texttt{crypto/}: Python cryptographic implementation for BabyJubJub ElGamal and tally utilities.
    \item \texttt{test/} and \texttt{tests/}: JavaScript/Hardhat and Python/Pytest verification suites.
\end{itemize}

This modularity enables independent testing of the smart-contract logic, the crypto primitive and the user-facing flow.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
    \node[draw, rounded corners=3pt, line width=0.4pt, text width=11.2cm, minimum height=1.2cm, align=center, inner xsep=7pt, fill=gray!8] (ui) at (0,0) {Frontend Layer: voter page, admin page, wallet interaction};
    \node[draw, rounded corners=3pt, line width=0.4pt, text width=11.2cm, minimum height=1.2cm, align=center, inner xsep=7pt, fill=gray!14] (crypto) at (0,-1.95) {Client Crypto Layer: ElGamal helpers, witness builder, Merkle audit library};
    \node[draw, rounded corners=3pt, line width=0.4pt, text width=11.2cm, minimum height=1.2cm, align=center, inner xsep=7pt, fill=gray!20] (chain) at (0,-3.9) {On-chain Layer: Voting (with voter registration), VoteVerifier, TallyVerifier};
    \node[draw, rounded corners=3pt, line width=0.4pt, text width=11.2cm, minimum height=1.2cm, align=center, inner xsep=7pt, fill=gray!26] (offchain) at (0,-5.85) {Off-chain Validation Layer: Hardhat tests, Python crypto tests};

    \draw[->, thick] (ui) -- (crypto);
    \draw[->, thick] (crypto) -- (chain);
    \draw[->, thick] (chain.south east) .. controls +(1.00,-0.55) and +(1.00,0.55) .. (offchain.north east);
    \draw[->, thick] (offchain.north west) .. controls +(-1.00,0.55) and +(-1.00,-0.55) .. (chain.south west);
    \draw[->, thick, dashed] (offchain.east) .. controls +(0.85,0.45) and +(0.85,-0.45) .. (ui.east);
\end{tikzpicture}
\caption{Implementation Layers and Verification Flows}
\label{fig:system-architecture}
\end{figure}

\noindent Figure~\ref{fig:system-architecture} depicts the architectural layering, principal data flows and components interactions. Solid arrows illustrate the operational workflows: runtime submission flows (Frontend $\rightarrow$ Crypto $\rightarrow$ Chain), and test-time validation exchange (Chain $\leftrightarrow$ Off-chain Validation). The dashed return arrow, by contrast, represents solely a development feedback loop and not carrying any on-chain protocol message.

\section{Smart Contract Design}

\subsection{Core Contract Responsibilities}

The voting contract is responsible for:
\begin{itemize}
    \item election initialization and candidate registration;
    \item lifecycle transition enforcement;
    \item vote acceptance conditioned on proof verification and voter eligibility;
    \item finalizing the tally based upon verification of proofs and consistency;
    \item Merkle-root publication for post-election audit.
\end{itemize}

No external voter/enforcer registry required as all the voters and enforcers enforce one vote themselves within their contracts: Voter, Voting. On-chain Groth16 verification method implemented by Verifier contracts.

\subsection{On-Chain Data Model}

Some of the main data-structures that are stored on chain are candidate details, voter's ballot information (i.e., commitment and cipher-text digest) for each voter, Commitments list etc., and continued output of the point-form tally results (\texttt{tallyResultPointX}/\texttt{tallyResultPointY}). The ciphertext payload itself are also emitted as events' values for saving on storage costs, yet maintaining the ability to reconstruct off-chain.

\subsection{Lifecycle Safeguards}

The contract modifiers implement the role/state restrictions. Forbidden phase transitions, unauthorised calls, double voting, and malformed proof submissions are rejected prior to state mutation. The only intentional exception is \texttt{updateMerkleRoot}: it continues being admin-callable until status \texttt{Tallied}. This governance tradeoff and its impact on assurances are discussed in Section~5.8 (P2: Merkle root mutability).

Table~\ref{tab:module-responsibility} summarizes how major modules split responsibilities.

\begin{table}[htbp]
\centering
\small
\caption{Module Responsibility Summary}
\label{tab:module-responsibility}
\begin{tabular}{L{3.0cm} L{4.0cm} L{5.8cm}}
\toprule
Module & Primary Input & Responsibility \\
\midrule
Voting contract & Proof data, commitments, tally results & Manage election lifecycle; verify proofs and enforce state transitions. \\
Vote circuit & Candidate ID, salt, ciphertext coordinates & Prove ballot validity and commitment-ciphertext binding. \\
Tally circuit & Aggregated ciphertexts, secret key & Prove correct decryption and vote-total consistency. \\
Voter frontend & Candidate selection, wallet signature & Encrypt ballot, generate vote proof, store receipt. \\
Admin frontend & Encrypted vote events, admin secret key & Aggregate/decrypt tally, generate ZKP3, build/export audit bundle. \\
\bottomrule
\end{tabular}
\end{table}

\section{Circuit Implementation}

\subsection{Vote Circuit}

The vote circuit integrates Poseidon commitment checks, candidate range checks, one-hot encryption consistency checks, and ciphertext-hash binding. Candidate range enforcement is implemented with \texttt{LessThan(8)}, so candidate indices are bounded to at most 255. In this repository, the instantiated configuration is fixed to \texttt{N=2}.

\subsection{Tally Circuit}

The tally circuit verifies key ownership relation and per-candidate decryption correctness for aggregated ciphertexts. For total consistency, it enforces equality at the curve-point level: the sum of public result points must equal \texttt{totalVotes} $\cdot G$. This is implemented with \texttt{Num2Bits(32)} for \texttt{totalVotes} encoding, which implies an upper bound of approximately $2^{32}-1$ on directly representable totals in-circuit. Public signals are exported and consumed by the Solidity verifier. Because consistency is point-based, strict integer-to-point binding is not fully enforced by the circuit alone; it would require additional contract-level constraints that remain unimplemented (see Section~5.8 for the corresponding P1 gap analysis).

\subsection{Artifact Pipeline}

The proof is generated following the normal Groth16 flow: circuit compiling, trusted setup artifacts, witness building and proving, and verifier calldata export \cite{circomdocs,snarkjs}. Verification is performed on the contract side using a pair-checking pre-compile.

\section{Frontend Engineering Workflow}

\subsection{Voter Interface}

The voter page executes the following sequence:
\begin{enumerate}
    \item initialize cryptographic runtime and contract connections;
    \item fetch candidate list and election state;
    \item on candidate selection, generate ciphertexts and commitment;
    \item generate vote proof and prepare calldata;
    \item submit transaction through wallet;
    \item persist local receipt for later anonymous verification.
\end{enumerate}

\subsection{Administrator Interface}

The administrator page performs:
\begin{enumerate}
    \item encrypted vote event retrieval;
    \item ciphertext reconstruction and integrity filtering by recomputing Poseidon \texttt{ciphertextHash} (mismatched entries are dropped);
    \item homomorphic aggregation, followed by decryption and candidate result extraction;
    \item tally proof generation (browser-side WASM in current implementation) and contract submission;
    \item deterministic commitment Merkle-tree construction after sorting by \texttt{blockNumber}, then \texttt{transactionIndex}, then \texttt{logIndex};
    \item audit-bundle export and local cache.
\end{enumerate}

\subsection{Audit Utility Module}

The audit utility in \texttt{frontend/lib/audit.js} centralizes:
\begin{itemize}
    \item bytes32 normalization;
    \item Merkle tree construction with duplicate-last odd handling;
    \item Merkle inclusion proof generation and path verification;
    \item election-key namespace generation for local storage isolation.
\end{itemize}

This separation reduces logic duplication and improves testability.

\section{Testing and Validation Strategy}

\subsection{Contract-Level Tests}

The Hardhat tests check the behaviour of roles controls, transitions between states, attempted vote creation (valid and invalid), tally submitter when using Groth16 proofs and end to end scenario such as generating real proofs and invoking a verifier. Measures have been taken in order not to increase too much the execution time of the test suite: costly Groth16 proofs are produced only once, during a common setup hook, and then recycled among the behavior-oriented tests.

\subsection{Cryptographic Unit Tests}

Pytest tests for the curve arithmetic invariants, key generation, encryption/decryption correctness, homomorphic addition, one-hot vote encoding semantics and bounded discrete-log solving.

\subsection{Engineering Observations}

The dual test stack gives coverage at the contract protocol level as well as the cryptographic primitive level, which gives high confidence in isolated module correctness. But, integration testing identified a few boundary cases that are notable to keep in mind during further development.

To begin with, the most brittle integration point is the mapping from circuit public signals to Solidity function parameters because snarkjs exports proof calldata in some particular encoding, any discrepancy of signal ordering between the circuit definition and the contract's desired \texttt{pubSignals} array results in a generic reject, not a descriptive per-signal error. This was addressed by careful alignment during circuit compilation, but requires maintenance discipline whenever circuits are recompiled.

Second, browser-side proof generation is resource-intensive. During development, memory pressure was observed on some low-memory devices and caused WASM instability. The voter interface should therefore include progress indicators and graceful error messages for proof generation failures.

Third, ElGamal tally decryption relies on bounded runtime discrete-log recovery (baby-step giant-step). In the current administrator workflow, the search bound is set dynamically to \texttt{totalVotes + 100}. If recovery falls outside this bound, the client raises an explicit error. Since each per-candidate tally is upper-bounded by \texttt{totalVotes}, this bound also upper-bounds per-candidate recovery in practice. The selected bound should still be documented and reviewed for larger-election deployment profiles.

\section{Deployment Notes}

Local deployment uses a Hardhat node and a Python-served static frontend. The deployment script (\texttt{scripts/deploy.js}) generates \texttt{frontend/config.js} automatically, which is required before the frontend pages can interact with the deployed contracts. In local mode, if \texttt{ELGAMAL\_SK} is not supplied, the script also generates a bootstrap administrator secret key and exposes it to the admin page as a visible hint; the administrator must still explicitly load that key before tallying.

For testnet or mainnet deployment, the following additional steps apply:
\begin{itemize}
    \item For non-local deployment, either provide \texttt{ELGAMAL\_SK} in \texttt{.env} or provide the public key coordinates \texttt{ELGAMAL\_PK\_X} and \texttt{ELGAMAL\_PK\_Y}. The current deployment script intentionally has no insecure fallback administrator key.
    \item Configure \texttt{.env} with the deployer private key and the target network RPC URL before running the deploy script.
    \item In the \texttt{frontend/zk/} directory, the built WASM witness generators and the \texttt{.zkey} proving keys of the deployed verifier contracts must be present.
    \item MetaMask should be set up on the voter's and administrator's browser with the proper network (the chain ID should match the deployment target).
    \item The trusted setup lineage (the \texttt{.ptau} that was used in key generation and circuit-specific \texttt{.zkey} files) must match the deployed verifiers. Recompiling circuits without redeploying verifiers will cause all proof submissions to be rejected.
\end{itemize}

The working dependencies are recorded in the repository's README, as well as in the deployment script.

\section{Chapter Summary}

This chapter turned the protocol design into actual code modules and execution traces. The main result is implementational traceability, i.e., all major protocol claims from Chapter~3 are linked with concrete contract logic, circuit constraints, frontend operation, and automated testing. Chapter~5 determines if it fulfills the above criteria, reports the performance attributes, and identifies remaining security vulnerabilities.

\endgroup

