\chapter{System Model and Methodology}
\begingroup
\justifying
\setlength{\parindent}{0pt}
\setstretch{1.5}
\setlength{\parskip}{0.5\baselineskip}
\titlespacing{\chapter}{0pt}{0pt}{0pt}
\titlespacing{\section}{0pt}{0pt}{0pt}

\section{Protocol Roles}

The protocol defines three distinct roles, each with a bounded set of privileges and responsibilities.

\textbf{Administrator.} The administrator deploys the election contracts, configures the candidate set, and maintains the voter whitelist. This role retains sole custody of the ElGamal private key $\mathit{sk}$ used to decrypt the aggregated tally. After the voting phase concludes, the administrator performs off-chain homomorphic aggregation, computes the decrypted result points, and submits the tally along with a Groth16 proof (ZKP3) to the contract. The semi-trusted assumption governing this role is formalised in Section~3.2.

\textbf{Voter.} Any Ethereum address that appears on the administrator-managed whitelist is eligible to cast exactly one ballot. The voter encrypts a one-hot candidate vector under the election public key, generates a zero-knowledge vote proof (ZKP1/ZKP2), and submits the encrypted ballot together with the proof to the contract. Upon acceptance, the voter retains a local receipt for subsequent audit verification (Section~3.9).

\textbf{Public verifier.} A public verifier is any party that can read on-chain state without holding privileged credentials. Public verifiers may observe lifecycle transitions, inspect Groth16 verification outcomes, replay the homomorphic tally from emitted ciphertext events, and execute the anonymous verification procedure described in Section~3.9 using the published audit bundle. This role embodies the system's transparency guarantee: all assurance properties except key custody are independently auditable from public data alone.

\section{Threat Model}

The model considers malicious voters and passive blockchain observers. For administrator behaviour, the prototype assumes a \emph{semi-trusted administrator}: the administrator is expected to follow the protocol, while single-key control remains an explicit trust assumption rather than a cryptographically enforced guarantee. The following attacks are explicitly targeted:
\begin{itemize}
    \item invalid ballot shape (non-one-hot encryption);
    \item out-of-range candidate encoding;
    \item replay or duplicate voting attempts;
    \item fabricated tally claims unsupported by valid proof.
\end{itemize}

The following are not fully solved in the current prototype: coercion resistance, malware-compromised browser clients, threshold key governance, fully adversarial administrator behavior (for example refusing to tally or submitting inconsistent off-chain artifacts), and per-candidate integer-to-result-point binding at contract layer (ZKP3 constrains curve-point correctness, while integer mapping remains client-verifiable rather than contract-enforced). Table~\ref{tab:threat-matrix} structures the full attack surface considered by this model.

\begin{table}[htbp]
\centering
\small
\caption{Threat Model Mapping Matrix}
\label{tab:threat-matrix}
\begin{tabular}{L{2.0cm} L{2.6cm} L{2.2cm} L{2.8cm} L{2.8cm}}
\toprule
Attacker & Attack method & Target property & Control mechanism & Residual risk \\
\midrule
Malicious voter & Submit non-one-hot or out-of-range encoding & Tally integrity & ZKP1/2 + contract range checks & Compromised client runtime can still submit seemingly valid inputs \\
Malicious voter & Replay/duplicate voting & One-person-one-vote & Registry whitelist + hasVoted gate & Credential theft/abuse is outside this model \\
Semi-trusted administrator & Infer individual ballot content & Ballot secrecy & ElGamal encryption + aggregate-only decryption + ZKP3 & Single-key custody remains a trust assumption \\
Adversarial administrator & Refuse tally or submit inconsistent off-chain artifacts & Availability and auditability & On-chain state machine + proof gates & Denial-of-service and off-chain artifact inconsistency not fully eliminated \\
Adversarial administrator & Alter per-candidate integer distribution (same total) & Per-candidate tally authenticity & Client-side point-to-integer recheck & Contract layer does not strongly bind integers to result points (see Section~5.8, \textbf{[P1]}) \\
\bottomrule
\end{tabular}
\end{table}

\section{Protocol State Machine}

Election lifecycle states are:
\begin{equation}
S = \{\texttt{Created},\ \texttt{Active},\ \texttt{Ended},\ \texttt{Tallied}\}
\end{equation}

Allowed transitions are:
\begin{equation}
\texttt{Created} \rightarrow \texttt{Active} \rightarrow \texttt{Ended} \rightarrow \texttt{Tallied}
\end{equation}

This deterministic state machine ensures that vote submission only occurs in the active phase and tally finalization only occurs after the election is ended.

Table~\ref{tab:state-transition-guards} makes the transition guards explicit at protocol level.

\begin{table}[htbp]
\centering
\small
\caption{State Transition Guards and Rejection Rules}
\label{tab:state-transition-guards}
\begin{tabular}{L{2.8cm} L{2.2cm} L{5.8cm}}
\toprule
Transition & Actor & Guard and rejection behaviour \\
\midrule
\texttt{Created} $\rightarrow$ \texttt{Active} & Administrator & \texttt{startElection()} requires candidate count equals circuit cardinality; otherwise contract reverts. \\
\texttt{Active} $\rightarrow$ \texttt{Ended} & Administrator & \texttt{endElection()} is accepted only in active state; otherwise contract reverts. \\
\texttt{Ended} $\rightarrow$ \texttt{Tallied} & Administrator & \texttt{updateTallyResults()} requires valid ZKP3, matching public key signals, and total-vote consistency; otherwise contract reverts. \\
\bottomrule
\end{tabular}
\end{table}

Figure~\ref{fig:protocol-flow} summarises the complete interaction sequence among administrator, voters, and smart contracts.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
    \node[draw, rounded corners, minimum width=8.6cm, minimum height=0.95cm] (a) at (0,0) {1) Setup: admin deploys contracts and publishes ElGamal public key};
    \node[draw, rounded corners, minimum width=8.6cm, minimum height=0.95cm] (b) at (0,-1.5) {2) Vote casting: voter encrypts one-hot ballot and generates vote proof};
    \node[draw, rounded corners, minimum width=8.6cm, minimum height=0.95cm] (c) at (0,-3.0) {3) On-chain verification: contract checks Groth16 proof and records commitment};
    \node[draw, rounded corners, minimum width=8.6cm, minimum height=0.95cm] (d) at (0,-4.5) {4) Tally: admin aggregates ciphertexts, decrypts totals, submits ZKP3};
    \node[draw, rounded corners, minimum width=8.6cm, minimum height=0.95cm] (e) at (0,-6.0) {5) Audit: voter verifies receipt commitment against Merkle proof};

    \draw[->, thick] (a) -- (b);
    \draw[->, thick] (b) -- (c);
    \draw[->, thick] (c) -- (d);
    \draw[->, thick] (d) -- (e);
\end{tikzpicture}
\caption{End-to-End Protocol Flow}
\label{fig:protocol-flow}
\end{figure}

Figure~\ref{fig:protocol-sequence} provides a sequence-style view of the same workflow with explicit message directions.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[x=1cm,y=1cm]
    \node[draw, minimum width=2.8cm, minimum height=0.75cm] (voter) at (0,0) {Voter Client};
    \node[draw, minimum width=3.2cm, minimum height=0.75cm] (chain) at (5.0,0) {Voting Contract};
    \node[draw, minimum width=2.8cm, minimum height=0.75cm] (admin) at (10.0,0) {Admin Client};

    \draw[dashed] (0,-0.45) -- (0,-7.1);
    \draw[dashed] (5.0,-0.45) -- (5.0,-7.1);
    \draw[dashed] (10.0,-0.45) -- (10.0,-7.1);

    \draw[->, thick] (0,-1.0) -- (5.0,-1.0);
    \node[font=\scriptsize, align=center] at (2.5,-0.72) {castVote(commitment, ciphertextHash, ZKP1/2)};

    \draw[->, thick] (5.0,-1.8) -- (0,-1.8);
    \node[font=\scriptsize, align=center] at (2.5,-1.52) {transaction receipt (event confirms commitment on-chain)};

    \draw[->, thick] (10.0,-2.8) -- (5.0,-2.8);
    \node[font=\scriptsize, align=center] at (7.5,-2.52) {query encrypted vote events};

    \draw[->, thick] (5.0,-3.6) -- (10.0,-3.6);
    \node[font=\scriptsize, align=center] at (7.5,-3.32) {event logs for homomorphic aggregation};

    \draw[->, thick] (10.0,-4.8) -- (5.0,-4.8);
    \node[font=\scriptsize, align=center] at (7.5,-4.52) {updateTallyResults(results, root, ZKP3)};

    \draw[->, thick] (5.0,-5.8) -- (0,-5.8);
    \node[font=\scriptsize, align=center] at (2.5,-5.52) {public tally + merkleRoot for audit verification};
\end{tikzpicture}
\caption{Protocol Sequence Diagram for Voting and Tally}
\label{fig:protocol-sequence}
\end{figure}

\section{Ballot Encoding and Encryption}

Let $N$ be the number of candidates and $\mathit{candidateId} \in \{0,1,\dots,N-1\}$ be the selected candidate index. A ballot is encoded as one-hot vector
$\mathbf{m}=(m_0,\dots,m_{N-1})$ where $m_{\mathit{candidateId}}=1$ and all other entries are $0$.

For each candidate position $j$, ElGamal ciphertext is generated on BabyJubJub:
\begin{equation}
C1_j = r_j \cdot G, \quad C2_j = m_j \cdot G + r_j \cdot \mathit{PK}
\end{equation}
where $r_j$ is sampled uniformly from the BabyJubJub scalar field, $G$ is generator, and $\mathit{PK}$ is administrator public key. We write $C1_j=(\mathit{c1X}_j,\mathit{c1Y}_j)$ and $C2_j=(\mathit{c2X}_j,\mathit{c2Y}_j)$ for affine coordinates, which are used by the vote-proof witness and ciphertext binding hash. The notation is general in $N$, while the current deployed prototype instantiates $N=2$ and enforces this cardinality at contract layer.

\section{Commitment and Ciphertext Binding}

For private random salt $\mathit{salt}$, commitment is defined as:
\begin{equation}
\mathit{commitment} = \mathsf{Poseidon}(\mathit{candidateId},\, \mathit{salt})
\end{equation}

Ciphertext binding hash is defined over all ciphertext coordinates:
\begin{equation}
\begin{split}
H_{ct} = \mathsf{Poseidon}(&\mathit{c1X}_0,\,\mathit{c1Y}_0,\,\mathit{c2X}_0,\,\mathit{c2Y}_0,\,\dots,\\
                  &\mathit{c1X}_{N-1},\,\mathit{c1Y}_{N-1},\,\mathit{c2X}_{N-1},\,\mathit{c2Y}_{N-1})
\end{split}
\end{equation}

The vote proof uses $(\mathit{commitment}, H_{ct}, \mathit{PK}_x, \mathit{PK}_y)$ as public signals and proves consistency between commitment, candidate index, and encrypted one-hot structure.
Here, $\mathit{PK}_x, \mathit{PK}_y$ denote affine coordinates of $\mathit{PK}$.
In the reference tally client, each emitted ciphertext payload is re-hashed off-chain and matched against the on-chain vote record $(\mathit{commitment}, \mathit{ciphertextHash})$ before aggregation; mismatched entries are excluded from the tally set. This is a client-side integrity requirement rather than a contract-enforced invariant.

\section{Vote Proof Constraints}

The vote circuit takes private inputs $(\mathit{candidateId},\, \mathit{salt},\, r_j,\, \mathit{c1X}_j,\, \mathit{c1Y}_j,\, \mathit{c2X}_j,\, \mathit{c2Y}_j)$ for $j \in \{0,\dots,N{-}1\}$, and exposes four public signals $(\mathit{commitment},\, H_{ct},\, \mathit{PK}_x,\, \mathit{PK}_y)$. It enforces the following constraint families.
\begin{enumerate}
    \item \textbf{Commitment consistency.} The circuit recomputes $\mathsf{Poseidon}(\mathit{candidateId},\,\mathit{salt})$ from private inputs and constrains the result to equal the public signal $\mathit{commitment}$. This binds the on-chain commitment to a specific candidate choice without disclosing it.
    \item \textbf{Candidate range.} The circuit asserts $0 \le \mathit{candidateId} < N$, where $N$ is the candidate cardinality hardcoded at compilation time. Any index outside this range causes the constraint system to be unsatisfiable.
    \item \textbf{One-hot encryption.} For each slot $j \in \{0,\dots,N{-}1\}$, the circuit verifies that the ElGamal ciphertext encodes $m_j = 1$ when $j = \mathit{candidateId}$ and $m_j = 0$ otherwise. Together with the range constraint, this guarantees that every accepted ballot assigns exactly one vote to exactly one candidate.
    \item \textbf{Ciphertext binding.} The circuit hashes all $4N$ ciphertext coordinates $(\mathit{c1X}_0,\,\mathit{c1Y}_0,\,\mathit{c2X}_0,\,\mathit{c2Y}_0,\,\dots,\,\mathit{c1X}_{N-1},\,\mathit{c1Y}_{N-1},\,\mathit{c2X}_{N-1},\,\mathit{c2Y}_{N-1})$ with Poseidon and constrains the digest to equal the public signal $H_{ct}$. Because $H_{ct}$ is recorded on-chain alongside the commitment, any post-proof substitution of ciphertext components is detectable.
\end{enumerate}

When the proof is accepted by the on-chain Groth16 verifier, these four constraints are satisfied simultaneously. The contract therefore treats the accompanying ballot as a well-formed, in-range, single-candidate vote whose encrypted payload is irrevocably bound to the published commitment.

\section{Homomorphic Tally and Tally Proof}

The additive homomorphism of ElGamal encryption permits the computation of aggregate vote totals without decrypting individual ballots. For each candidate position $j$, the aggregated ciphertext is obtained by elliptic-curve point addition over all $V$ cast ballots:
\begin{equation}
\bar{C1}_j = \sum_{k=1}^{V} C1_{k,j}, \quad \bar{C2}_j = \sum_{k=1}^{V} C2_{k,j}
\end{equation}

Because each individual ciphertext encrypts either $0 \cdot G$ or $1 \cdot G$ in the one-hot encoding, the aggregate satisfies:
\begin{equation}
\bar{C2}_j - \mathit{sk} \cdot \bar{C1}_j = \left(\sum_{k=1}^{V} m_{k,j}\right) \cdot G = n_j \cdot G
\end{equation}
where $n_j$ denotes the integer vote count for candidate $j$. The administrator decrypts the aggregated ciphertext by computing the result point for each candidate:
\begin{equation}
R_j = \bar{C2}_j - \mathit{sk} \cdot \bar{C1}_j
\end{equation}

Recovery of the integer $n_j$ from the curve point $R_j = n_j \cdot G$ requires a bounded discrete-logarithm search. Since $n_j \le V$ and the total number of voters is practically bounded, the search remains computationally feasible. The administrator submits the recovered integer results to the contract together with a Groth16 proof of decryption correctness.

The tally circuit takes a single private input $\mathit{sk}$ and exposes all remaining values as public signals: $(\mathit{PK}_x,\, \mathit{PK}_y,\, \bar{C1}^{X}_j,\, \bar{C1}^{Y}_j,\, \bar{C2}^{X}_j,\, \bar{C2}^{Y}_j,\, R^{X}_j,\, R^{Y}_j,\, \mathit{totalVotes})$ for $j \in \{0,\dots,N{-}1\}$. It enforces three constraint families.
\begin{enumerate}
    \item \textbf{Key ownership.} The circuit verifies $\mathit{PK} = \mathit{sk} \cdot G$, confirming that the prover holds the secret key corresponding to the on-chain election public key.
    \item \textbf{Decryption correctness.} For each candidate $j$, the circuit constrains $R_j = \bar{C2}_j - \mathit{sk} \cdot \bar{C1}_j$, ensuring that each result point is the honest decryption of the corresponding aggregated ciphertext.
    \item \textbf{Sum consistency.} The circuit enforces $\sum_{j=0}^{N-1} R_j = \mathit{totalVotes} \cdot G$ in curve-point arithmetic, confirming that the individual result points collectively account for the declared total number of votes.
\end{enumerate}

Before invoking the Groth16 verifier, the contract performs two pre-checks: the public-key coordinates in the proof's public signals must match the on-chain election key, and the $\mathit{totalVotes}$ signal must equal the contract's internal vote counter. These guards ensure that the proof pertains to the correct election instance and voter population. If the proof is accepted, the contract stores the submitted integer results and the Merkle root, then transitions the election to the \texttt{Tallied} state.

The circuit binds correctness at the level of curve points rather than integers. Per-candidate integer-to-point correspondence ($n_j \cdot G = R_j$) is independently verifiable by any auditor who recomputes the scalar multiplication, but this check is not enforced by contract-level constraints. This design boundary and its implications are discussed further in Section~5.8.

\section{Merkle Audit Construction}

The Merkle audit tree provides a compact digest of all $V$ ballot commitments, enabling $O(\log V)$ inclusion verification for individual voters. The tree takes as input the ordered sequence of on-chain commitments and produces a single root hash that is stored on-chain upon tally finalisation.

Leaf nodes are the Poseidon commitment values themselves, used directly without additional hashing:
\begin{equation}
\mathit{leaf}_i = \mathit{commitment}_i
\end{equation}

Internal nodes are computed using Keccak256 over the concatenation of their two children:
\begin{equation}
\mathit{node} = \mathsf{Keccak256}(\mathit{left} \parallel \mathit{right})
\end{equation}

This hybrid construction---Poseidon leaves produced by the vote circuit and Keccak256 internal nodes for EVM-native compatibility---balances arithmetic-circuit efficiency at the leaf level with gas-efficient on-chain verification at intermediate levels.

The tree construction observes three deterministic rules that together guarantee reproducibility.

\textbf{Canonical ordering.} Commitments are sorted lexicographically by $(\texttt{blockNumber},\, \texttt{transactionIndex},\, \texttt{logIndex})$. This total ordering ensures that any implementation processing the same set of on-chain events will compute an identical root.

\textbf{Odd-layer padding.} When a tree layer contains an odd number of nodes, the final node is duplicated to serve as its own sibling, thereby completing the binary pairing for that layer.

\textbf{Root immutability.} Once the election transitions to the \texttt{Tallied} state, a contract-level state guard rejects any subsequent call to update the \texttt{merkleRoot}. The audit digest is therefore fixed on-chain from the point of finalisation onward.

Proving the inclusion of any single commitment requires only the $O(\log V)$ sibling hashes along the leaf-to-root path. Voters combine this Merkle proof with their locally stored receipt to execute the anonymous verification procedure described in Section~3.9.

\section{Anonymous Verification Procedure}

This procedure enables a voter to confirm that their ballot commitment was included in the finalised audit set, without revealing their identity or ballot content during the verification step. The voter retains a local receipt containing five fields: candidate identifier, salt, commitment, transaction hash, and election identifier.

Verification proceeds through four sequential checks:
\begin{enumerate}
    \item Recompute the commitment as $\mathsf{Poseidon}(\mathit{candidateId},\, \mathit{salt})$ from the receipt fields and confirm that the result matches the stored commitment value.
    \item Locate the entry in the audit bundle whose commitment and transaction hash both match the receipt.
    \item Verify the Merkle inclusion path from the matched leaf to the bundle root using the sibling hashes provided in the bundle entry.
    \item Accept if and only if the Merkle path is valid, the bundle root equals the on-chain \texttt{merkleRoot}, and the election identifier in the bundle metadata matches the receipt.
\end{enumerate}

Because the procedure relies on commitment and transaction hash rather than the voter's address, no address-based lookup is required. However, on-chain vote events still record the sender address alongside the commitment; this linkage is visible at the chain layer and falls outside the scope of this verification-flow abstraction.

Table~\ref{tab:audit-artifacts} summarises the artifacts involved in anonymous verification.

\begin{table}[htbp]
\centering
\small
\caption{Audit Artifacts and Their Roles}
\label{tab:audit-artifacts}
\begin{tabular}{L{3.2cm} L{3.8cm} L{5.8cm}}
\toprule
Artifact & Source & Verification Role \\
\midrule
Receipt JSON & Voter local storage & Provides candidate ID, salt, and commitment to recompute commitment consistency. \\
Commitment leaf & On-chain vote event & Represents one ballot commitment in Merkle construction. \\
Proof path & Audit bundle entry & Enables logarithmic-size inclusion verification against bundle root. \\
Merkle root & Tally publication / bundle metadata & Serves as commitment set digest to validate inclusion outcome. \\
Election ID & Bundle + receipt metadata & Prevents cross-election replay or accidental bundle mismatch. \\
\bottomrule
\end{tabular}
\end{table}

\endgroup

