\chapter{Literature Review}
\begingroup
\justifying
\setlength{\parindent}{0pt}
\setstretch{1.5}
\setlength{\parskip}{0.5\baselineskip}
\titlespacing{\chapter}{0pt}{0pt}{0pt}
\titlespacing{\section}{0pt}{0pt}{0pt}

Chapter~2 surveys related prior work on e-voting systems and analyzes the details of the cryptographic foundations adopted in this prototype.

It opens by defining the security requirements all reliable e-voting systems should satisfy. Centralised architectures suffer from inherent single points of failure and problematic trust assumptions. Purely blockchain-based systems address transparency but leave the privacy--verifiability gap unresolved. This dissertation positions itself at the intersection of these two extremes.

This chapter then introduces the main cryptographic constructions utilized in the implementation: ElGamal homomorphic encryption instantiated over BabyJubJub, Groth16 zero-knowledge proofs, and commitment-based Merkle trees that support efficient auditability.

\section{Security Requirements of E-Voting}

Any trustworthy e-voting system must satisfy four essential security requirements: eligibility control, ballot secrecy, individual verifiability, and universal verifiability.

Eligibility means only authorized voters can cast a vote and can only vote once. Ballot secrecy requires that no entity, including the administrator, be able to reveal any specific ballot. Individual verifiability allows every voter to verify if their own ballot was correctly counted. Universal verifiability means that any third-party observer may check whether the published result matches the set of cast ballots.

This dissertation addresses all four properties at the cryptographic level. Eligibility is enforced through an on-chain voter whitelist registry. Homomorphic encryption is adopted to achieve ballot secrecy, so that plaintext choices are never stored. Individual verifiability is provided by receipt-based Merkle inclusion proofs. Universal verifiability is enabled through on-chain Groth16 zero-knowledge proof verification of both vote validity and tally correctness.

\section{Centralised E-Voting Systems}

Conventional centralised systems handle the full election lifecycle through web services and database transactions. The approach has practical advantages: mature tooling, straightforward deployment, and low operational complexity. Several national deployments have adopted this model at scale.

Estonia's internet voting system, operational since 2005, represents the most mature large-scale deployment. Voters authenticate using national identity cards and may revise their ballot multiple times before polling closes. A security audit by Springall et al.\ in 2014 identified server-side vulnerabilities, inconsistent software validation, and an absence of independently verifiable evidence that tallied results matched the cast ballots \cite{springall2014}. The audit concluded that the system's assurances rest primarily on trust in operator infrastructure rather than on publicly verifiable cryptographic guarantees.

Helios, introduced by Adida in 2008, adopts a different philosophy \cite{adida2008}. It publishes encrypted ballots and decryption proofs openly, enabling independent tally verification without trusting the operator's claimed output. Helios has been used in real organisational elections, including internal votes of the International Association for Cryptologic Research. Its trust model, however, still depends on a central server to collect and publish ciphertexts; a compromised server can suppress or substitute ballots before they enter the public record.

These examples reveal a structural limitation that centralised architectures cannot resolve through engineering refinement alone:
\begin{itemize}
    \item privileged insiders can alter records or logs if access controls fail;
    \item replaying state transitions independently requires a full data export, which operators may restrict;
    \item dispute resolution ultimately falls back on institutional trust rather than any independently verifiable record.
\end{itemize}

The root cause is architectural: all integrity checks operate within the same trust boundary as the operators who control the system. Replacing that boundary with a tamper-evident, publicly verifiable record motivates the turn to blockchain infrastructure.

\section{Blockchain-Based Voting Approaches}

Blockchain infrastructures maintain immutable append-only logs with deterministic execution rules \cite{nakamoto2008,wood2014}. Election state transitions and event traces are transparently recorded. Smart contracts enforce lifecycle invariants directly, removing the hidden backend logic that centralised systems rely on.

Early work on blockchain-based voting focused on demonstrating feasibility rather than resolving privacy. McCorry, Shahandashti, and Hao published the first serious treatment of privacy-preserving smart-contract voting in 2017 \cite{mccorry2017}. Their protocol, implemented on Ethereum, uses public-key encryption to conceal individual ballots and ensures the administrator learns only the aggregate outcome. The scheme requires all voters to be online simultaneously during a commitment phase, however, and verification involves a multi-round interactive protocol. These interaction requirements limit practical applicability for general elections.

Hjalmarsson et al.\ examined the scalability and deployment characteristics of blockchain e-voting in 2018, implementing a prototype on Ethereum and measuring transaction costs and throughput under realistic conditions \cite{hjalmarsson2018}. Their results confirmed that on-chain lifecycle enforcement is technically feasible, but identified gas costs and block-rate constraints as binding limits for large electorates. Ballot content in their prototype is stored in plaintext, leaving ballot secrecy unresolved.

Transparency cuts both ways. If vote semantics appear in plaintext on-chain, confidentiality is permanently and irrecoverably lost. A privacy-preserving blockchain voting system must therefore pair ledger transparency for control flow with cryptographic concealment for ballot content. Existing proposals address this tension incompletely: systems that encrypt ballots off-chain and submit only ciphertexts still require a trusted party to attest to the decryption without a verifiable proof, so the tally claim rests on operator honesty rather than cryptographic evidence.

The gap is not a missing component but a missing integration: encrypting ballots, proving their validity without revealing their content, proving the correctness of the aggregate decryption, and enabling post-election inclusion auditing must all operate together in a single verifiable pipeline.

\section{Positioning of This Work}

This dissertation addresses that integration gap. It builds and evaluates a complete pipeline connecting browser-side proof generation, smart-contract verification, and post-election audit tooling in one deployable workflow.

The contribution has two concrete parts:
\begin{itemize}
    \item a working prototype demonstrating that homomorphic encryption, zero-knowledge proof verification, and Merkle auditability operate together under one architecture;
    \item an explicit assessment of which security goals are fully met, partially met, or still unresolved, providing a grounded basis for future hardening.
\end{itemize}

Table~\ref{tab:approach-comparison} summarises the most relevant architectural tradeoffs across the three categories of system examined in this chapter.

\begin{table}[htbp]
\centering
\small
\caption{Architectural Comparison Across Voting Approaches}
\label{tab:approach-comparison}
\begin{tabular}{L{3.0cm} L{2.9cm} L{2.9cm} L{3.9cm}}
\toprule
Criterion & Centralised E-Voting & Blockchain-Only Voting & This Work (Blockchain + HE + ZKP + Merkle) \\
\midrule
State transparency & Limited by operator logs & High for on-chain actions & High for control flow and published commitments \\
Ballot confidentiality & Depends on backend policy & Usually weak if plaintext on-chain & Stronger via encrypted ballots and commitment receipts \\
Tally verifiability & Operational trust heavy & Moderate unless cryptographic proofs added & High through explicit vote/tally proofs \\
Voter-side inclusion check & Rare or account-linked & Possible but privacy-sensitive & Receipt-based Merkle inclusion workflow with reduced address-linkage during verification \\
Operational complexity & Lower & Moderate & Higher, but with stronger assurance properties \\
\bottomrule
\end{tabular}
\end{table}

The following sections develop the cryptographic mechanisms that underpin this integration.

\section{Homomorphic Encryption for Private Tally}

Two principal homomorphic encryption schemes appear in e-voting literature. Paillier encryption \cite{paillier1999} supports additive homomorphism over integer plaintexts directly, without point encoding, and has been adopted in systems such as Helios \cite{adida2008}. Its ciphertext arithmetic operates modulo a composite integer, however, which does not map efficiently to the finite-field arithmetic that zk-SNARK circuits require. Embedding Paillier operations inside a Groth16 circuit would require non-native field arithmetic, substantially increasing constraint counts.

ElGamal encryption on elliptic-curve discrete-log groups supports additive homomorphism over encoded messages \cite{elgamal1985,hankerson2004ecc}. This work instantiates ElGamal on BabyJubJub \cite{eip2494}, which is defined over the BN254 scalar field and therefore admits native arithmetic in Groth16 circuits without field-conversion overhead. Each ballot is encoded as a one-hot vector; component-wise aggregation of ciphertexts gives encrypted candidate totals. Decrypting only the aggregates avoids exposing individual ballots.

Two practical advantages follow. Ciphertext aggregation scales linearly with vote count using simple point additions; recovering final integer totals from decrypted points is a separate off-chain discrete-log step. Privacy holds as long as secret-key exposure and plaintext side channels stay controlled---a weaker guarantee than threshold decryption, but achievable without ceremony infrastructure.

\section{Zero-Knowledge Proofs in Voting}

A zero-knowledge proof (ZKP) is a cryptographic protocol in which a prover convinces a verifier that a statement is true without disclosing any information beyond the statement's truth itself \cite{goldwasser1989}. In e-voting, this capability is foundational: a voter must prove that their ballot is correctly formed---encoding exactly one valid candidate choice---without revealing which candidate was selected; an administrator must prove that a tally is correctly decrypted without revealing the secret key. ZKPs therefore allow public verifiability and ballot privacy to coexist.

This dissertation uses Groth16, a zk-SNARK (Succinct Non-interactive ARgument of Knowledge), because its pairing-check cost is fixed with respect to circuit-constraint count and proof size is constant at three elliptic-curve points \cite{groth2016}. Total verifier gas still includes a smaller public-input term, but does not scale with full circuit size. Circom and snarkjs provide the implementation pipeline from circuit specification to deployable Solidity verifier contracts \cite{circomdocs,snarkjs}.

\subsection{The Problem ZKPs Address}

Many non-zero-knowledge proof workflows require witness (secret data) exposure to the verifier. In e-voting this is unacceptable: revealing a candidate index to prove ballot validity destroys ballot secrecy; revealing a decryption key to prove tally correctness collapses key governance.

ZKPs separate the \emph{claim} from the \emph{evidence}. Let $x$ be a public statement and $w$ be a private witness. Define a relation $\mathcal{R}$ such that $(x, w) \in \mathcal{R}$ means ``$w$ is a valid witness for $x$''. A ZKP system allows a prover who knows $w$ to produce a short artifact $\pi$ that convinces any verifier that $(x, w) \in \mathcal{R}$, without $\pi$ leaking any information about $w$.

In this project, two instances of this pattern apply:
\begin{itemize}
    \item \textbf{Vote proof}: the public statement and witness are
    \begin{align*}
    x &= (\mathit{commitment},\; H_{ct},\; PK_x,\; PK_y) \\
    w &= (\mathit{candidateId},\; \mathit{salt},\; r_0, \dots, r_{N-1},\\
       &\quad \mathit{c1X}_0,\; \mathit{c1Y}_0,\; \mathit{c2X}_0,\; \mathit{c2Y}_0,\; \dots,\; \mathit{c1X}_{N-1},\; \mathit{c1Y}_{N-1},\; \mathit{c2X}_{N-1},\; \mathit{c2Y}_{N-1})
    \end{align*}
    The relation $\mathcal{R}$ requires that: (i) $\mathit{commitment} = \mathsf{Poseidon}(\mathit{candidateId}, \mathit{salt})$, (ii) $0 \le \mathit{candidateId} < N$, (iii) each ciphertext tuple $(\mathit{c1X}_j,\mathit{c1Y}_j,\mathit{c2X}_j,\mathit{c2Y}_j)$ is a valid ElGamal encryption of one-hot bit $m_j$ where $m_j=1$ iff $j=\mathit{candidateId}$, and (iv) $H_{ct}$ matches the Poseidon hash of all ciphertext coordinates.
    \item \textbf{Tally proof}: $w = \mathit{sk}$, with public statement
    \begin{equation*}
        \begin{split}
        x = (&PK_x,\; PK_y,\; \bar{C1}^{X}_0,\; \bar{C1}^{Y}_0,\; \dots,\; \bar{C1}^{X}_{N-1},\; \bar{C1}^{Y}_{N-1},\\
             &\bar{C2}^{X}_0,\; \bar{C2}^{Y}_0,\; \dots,\; \bar{C2}^{X}_{N-1},\; \bar{C2}^{Y}_{N-1},\\
             &R^{X}_0,\; R^{Y}_0,\; \dots,\; R^{X}_{N-1},\; R^{Y}_{N-1},\; \mathit{totalVotes})
        \end{split}
    \end{equation*}
    where $R_j = n_j \cdot G$ is the result point for candidate $j$, $R^{X}_j$, $R^{Y}_j$ are the coordinates of $R_j$, $n_j$ is the corresponding integer vote count, and $\mathit{totalVotes}$ is the expected sum of all candidate totals (corresponding to $V$ in Chapter 1). Here, $PK_x, PK_y$ denote affine coordinates of $\mathit{PK}$. Likewise, $\bar{C1}^{X}_j,\bar{C1}^{Y}_j$ and $\bar{C2}^{X}_j,\bar{C2}^{Y}_j$ denote affine coordinates of aggregated points $\bar{C1}_j,\bar{C2}_j$.
\end{itemize}

\subsection{Formal Security Properties}

A ZKP system must satisfy three properties \cite{goldwasser1989}:
\begin{enumerate}
    \item \textbf{Completeness}: If $(x, w) \in \mathcal{R}$, then an honest prover holding $w$ always produces a proof $\pi$ that the verifier accepts. This guarantees that valid ballots and correct tallies are never falsely rejected.
    \item \textbf{Soundness}: If no valid witness $w$ exists for statement $x$, then no computationally bounded adversary can produce an accepted proof except with negligible probability. This prevents fabricated ballots or incorrect tally claims from being accepted on-chain.
    \item \textbf{Zero-Knowledge}: A simulator can produce proofs with the same distribution as real proofs without knowing $w$. This formalizes that $\pi$ reveals nothing about $w$ beyond the existence of a valid witness.
\end{enumerate}

In a zk-SNARK, soundness is computational (relying on cryptographic hardness assumptions), and non-interactivity is achieved through a structured reference string (SRS) produced in a trusted setup. The SRS provides the cryptographic parameters needed for proving and verification without any round-trip interaction.

\subsection{Arithmetic Circuits and R1CS}

Any computation over a finite field $\mathbb{F}_p$ can be expressed as an arithmetic circuit of addition and multiplication gates. A circuit is \emph{satisfiable} if there exists an assignment to all wires (public inputs, private witnesses, and intermediate values) that makes every gate equation hold. This satisfiability problem is encoded as a \emph{Rank-1 Constraint System} (R1CS).

An R1CS consists of $m$ constraints over a witness vector $\mathbf{s} \in \mathbb{F}_p^n$ (which includes public inputs, private witnesses, and intermediate wire values, with $s_0 = 1$ by convention):
\begin{equation}
    \langle \mathbf{a}_i,\, \mathbf{s} \rangle \;\cdot\; \langle \mathbf{b}_i,\, \mathbf{s} \rangle \;=\; \langle \mathbf{c}_i,\, \mathbf{s} \rangle, \qquad i = 1, \dots, m
\end{equation}
where $\mathbf{a}_i, \mathbf{b}_i, \mathbf{c}_i \in \mathbb{F}_p^n$ are sparse selector vectors derived from the circuit topology. Each multiplication gate contributes one constraint; additions and constant multiplications are absorbed into the selector vectors at no constraint cost.

In circom, the programmer writes constraints in a domain-specific language. The compiler translates them to an \texttt{.r1cs} file (the constraint matrices $A, B, C$) and a WASM witness generator. For this project, the compiled constraint counts are:
\begin{itemize}
    \item \texttt{vote\_proof}: 14{,}590 constraints, encoding commitment verification, candidate range checks, one-hot encryption validity, and ciphertext hash binding.
    \item \texttt{tally\_proof}: 16{,}786 constraints, encoding key-ownership verification, per-candidate decryption correctness, and total-vote sum consistency.
\end{itemize}

\subsection{QAP Transformation}

Groth16 operates on a polynomial encoding of the R1CS called a \emph{Quadratic Arithmetic Program} (QAP) \cite{gennaro2013}. For each constraint index $i$ and witness index $j$, define polynomials $u_j, v_j, w_j \in \mathbb{F}_p[X]$ by Lagrange interpolation through the point values $u_j(i) = a_{i,j}$, $v_j(i) = b_{i,j}$, $w_j(i) = c_{i,j}$.

Define
\begin{equation*}
A(X)=\sum_j s_j\,u_j(X), \qquad B(X)=\sum_j s_j\,v_j(X), \qquad C(X)=\sum_j s_j\,w_j(X).
\end{equation*}
At each constraint point $X=i$, interpolation gives $A(i)=\langle \mathbf{a}_i,\mathbf{s}\rangle$, $B(i)=\langle \mathbf{b}_i,\mathbf{s}\rangle$, and $C(i)=\langle \mathbf{c}_i,\mathbf{s}\rangle$. Therefore the R1CS equations imply
\begin{equation*}
A(i)\cdot B(i)-C(i)=0, \qquad i=1,\dots,m.
\end{equation*}
So the polynomial
\begin{equation}
\begin{split}
    P(X) \;=\; &\Bigl(\sum_j s_j\, u_j(X)\Bigr) \cdot \Bigl(\sum_j s_j\, v_j(X)\Bigr) \\
               &-\; \sum_j s_j\, w_j(X)
\end{split}
\end{equation}
satisfies $P(i)=0$ for all $i=1,\dots,m$. Since $T(X)=\prod_{i=1}^{m}(X-i)$ has exactly these roots, $T(X)$ divides $P(X)$, and equivalently there exists a quotient polynomial $H(X)$ such that:
\begin{equation}
    A(X) \cdot B(X) \;-\; C(X) \;=\; H(X) \cdot T(X)
\end{equation}
where $A, B, C$ are shorthand for the left, right, and output linear combinations of the witness polynomials. The prover's task is to commit to all polynomials evaluated at a secret point $\tau$, without knowing $\tau$ in the clear. This is achieved through the structured reference string produced in the trusted setup phase.

\subsection{The Groth16 Protocol}

Groth16 implements the QAP argument using bilinear pairings over elliptic curves. Let $\mathbb{G}_1, \mathbb{G}_2$ be groups of prime order $p$ with generators $G_1, G_2$, and let $e: \mathbb{G}_1 \times \mathbb{G}_2 \rightarrow \mathbb{G}_T$ be a non-degenerate bilinear pairing. The protocol has three phases.

\paragraph{Setup.}
A trusted party samples secrets $(\tau, \alpha, \beta, \gamma, \delta) \overset{\$}{\leftarrow} \mathbb{F}_p^5$ and computes two structured reference strings:
\begin{itemize}
    \item \textbf{Proving key} $\mathit{pk}_{\mathit{Groth}}$: contains $\{[\tau^i]_1\}$ for required powers, $\{[u_j(\tau)]_1\}$, $\{[v_j(\tau)]_2\}$, and blinded combinations---tens of thousands of elliptic-curve points for circuits of this scale (notation distinguished from ElGamal public key $\mathit{PK}$ and coordinates $PK_x,PK_y$).
    \item \textbf{Verification key} $\mathit{vk}$: contains $[\alpha]_1$, $[\beta]_2$, $[\gamma]_2$, $[\delta]_2$, and $\{[(\beta u_j + \alpha v_j + w_j)(\tau)/\gamma]_1\}$ for the public signal indices.
\end{itemize}

The secrets $(\tau, \alpha, \beta, \gamma, \delta)$ are then irrecoverably destroyed---these are the ``toxic waste'' of the trusted setup. If any value were leaked, an adversary could forge proofs for false statements without possessing a valid witness. This risk motivates the \emph{Powers of Tau} multi-party computation ceremony \cite{powersoftau2017}, in which many independent participants each contribute fresh randomness and prove they discarded their contribution. The final artifact is secure as long as at least one participant acted honestly.

\paragraph{Prove.}
Given the full witness $\mathbf{s}$ and the proving key, the prover samples random blinding scalars $\rho, \sigma \overset{\$}{\leftarrow} \mathbb{F}_p$ and computes a three-element proof $\pi = (\pi_A, \pi_B, \pi_C)$ (distinct from ElGamal encryption randomness $r_j$):
\begin{align}
    \pi_A &= \bigl[\alpha + A(\tau) + \rho\delta\bigr]_1 \\
    \pi_B &= \bigl[\beta + B(\tau) + \sigma\delta\bigr]_2 \\
    \pi_C &= \left[\frac{C_{\mathit{priv}}(\tau) + H(\tau)T(\tau)}{\delta}\right]_1 + \sigma \cdot \pi_A \\
          &\quad + \rho \cdot \bigl[\beta + B(\tau) + \sigma\delta\bigr]_1 - \rho\sigma \cdot [\delta]_1 \notag
\end{align}
where $[\cdot]_k$ denotes the elliptic-curve point $(\cdot) \cdot G_k$, and $C_{\mathit{priv}}$ denotes the private-witness portion of the $C$ polynomial. The blinding factors $\rho$ and $\sigma$ enforce Groth16's perfect zero-knowledge property: a simulator can generate proofs with the same distribution as real proofs without knowing the witness.

\paragraph{Verify.}
The verifier checks one bilinear pairing equation combining proof elements with verification-key elements:
\begin{equation}
    e(\pi_A,\; \pi_B) \;=\; e\bigl([\alpha]_1,\; [\beta]_2\bigr) \;\cdot\; e\!\left(\sum_{i=0}^{l} a_i \cdot \mathit{vk}_i,\; [\gamma]_2\right) \;\cdot\; e(\pi_C,\; [\delta]_2)
\end{equation}
where $l$ is the number of explicit public inputs (excluding the constant term), $\mathit{vk}_i = [(\beta u_i + \alpha v_i + w_i)(\tau)/\gamma]_1$ are precomputed verification-key elements, $a_0 = 1$ is the constant term, and $a_1,\dots,a_l$ are the public-input values. This single equation encodes the correctness of $A(\tau) \cdot B(\tau) = C(\tau) + H(\tau) \cdot T(\tau)$ through the algebraic properties of the pairing, without recomputing the polynomials at $\tau$. A proof is accepted if and only if this equation holds.

\subsection{On-Chain Verification Cost}

The BN254 curve (also called BN128) is natively supported by the EVM as a precompiled contract at address \texttt{0x08} for pairing computations. Under EIP-1108 \cite{eip1108}, the gas cost model is:
\begin{equation}
    \mathit{gas}_{\mathit{pairing}} = 45{,}000 + 34{,}000 \times k
\end{equation}
where $k$ is the number of pairing pairs. The Groth16 verification equation requires $k = 4$ pairs (corresponding to the four terms in the equation above), yielding $45{,}000 + 34{,}000 \times 4 = 181{,}000$ gas for the pairing check alone. Crucially, this pairing-check cost is \emph{independent of the number of circuit constraints}: a 14{,}590-constraint vote circuit and a 100{,}000-constraint circuit incur identical pairing gas. The public-input linear combination adds a smaller overhead proportional to the number of public inputs, not circuit constraints. This property makes Groth16 uniquely suited for complex ZKP computations that must be verified on-chain.

\subsection{Implementation Pipeline}

The implementation follows a four-stage workflow:
\begin{enumerate}
    \item \textbf{Circuit specification}: constraints are written in circom using library templates for BabyJubJub arithmetic, Poseidon hashing, and conditional logic. The compiler outputs \texttt{.r1cs} (constraint matrices $A, B, C$) and a WASM witness generator.
    \item \textbf{Trusted setup}: the Powers of Tau artifact (\texttt{pot16\_final.ptau}) provides a universal SRS supporting up to $2^{16}$ constraints. Circuit-specific phase-2 setup uses snarkjs to specialize the SRS to the particular R1CS, producing the proving key (\texttt{.zkey} file, about 7--10 MB per circuit in this implementation).
    \item \textbf{Proof generation}: given public inputs and private witnesses, the WASM generator constructs the full witness vector $\mathbf{s}$; snarkjs then computes $(\pi_A, \pi_B, \pi_C)$ using the proving key. In the current repository revision, local Node.js measurements place vote-proof generation at approximately 0.6--1.1 seconds for the instantiated $N=2$ circuit, while browser-side snarkjs WASM execution is typically similar or moderately slower depending on device hardware and caching behavior.
    \item \textbf{Verifier export}: snarkjs generates a Solidity contract that hard-codes the verification key and implements the pairing-check equation as direct EVM precompile calls, eliminating runtime parsing overhead.
\end{enumerate}

A key trust property of this pipeline is that the \emph{circuit code is public}: anyone can read \texttt{vote\_proof.circom} and \texttt{tally\_proof.circom}, recompile them, and verify that the resulting R1CS matches the corresponding \texttt{.zkey} artifacts when those artifacts are published with the deployment. This auditability---combined with the verifiable Powers of Tau ceremony transcript---means that trust in the system rests on the cryptographic hardness assumptions and the setup ceremony, not on the circuit author's integrity.

\subsection{ZKP Role in This Dissertation}

Three distinct proof instances are deployed, organised into two circuit files:
\begin{enumerate}
    \item \textbf{ZKP1 (commitment correctness)}: proves $\mathit{commitment} = \mathsf{Poseidon}(\mathit{candidateId}, \mathit{salt})$ without revealing $\mathit{candidateId}$ or $\mathit{salt}$.
    \item \textbf{ZKP2 (one-hot validity)}: proves that the submitted ElGamal ciphertexts encrypt precisely the one-hot vector $\mathbf{m}$ consistent with the committed candidate, without revealing $\mathbf{m}$ or the per-component randomness. Because ZKP1 and ZKP2 share witness variables (both depend on $\mathit{candidateId}$), they are compiled into a single \texttt{vote\_proof} circuit. A single proof simultaneously establishes both properties.
    \item \textbf{ZKP3 (tally decryption correctness)}: proves that for each candidate $j$, the published result point $R_j = n_j \cdot G$ equals the point obtained by decrypting the aggregated ciphertext under $\mathit{sk}$, and that $\mathit{sk}$ is consistent with the public key $\mathit{PK} = \mathit{sk} \cdot G$. The integer vote counts are derived off-chain from these curve points via discrete-log recovery, and are submitted to the contract separately; the circuit proves the correctness of the curve-point representations, not the integer values directly. This circuit is necessarily separate because it operates over the aggregated ciphertext state, which is only defined after all votes have been collected.
\end{enumerate}

The on-chain verifier contracts (\texttt{VoteVerifier.sol}, \texttt{TallyVerifier.sol}), generated automatically by snarkjs, each implement the pairing-check equation as a pure Solidity function. The main \texttt{Voting} contract calls these verifiers as a mandatory precondition for \texttt{castVote} and \texttt{updateTallyResults} respectively, enforcing the encoded circuit relations before state updates. Chapter~5 discusses remaining binding gaps that still require hardening for production assurance.

\section{Commitments, Hashing, and Merkle Auditability}

In this design, a commitment acts as a binding receipt for hidden ballot data. The voter computes a hash from candidate ID and salt, then publishes only the digest. Later edits to either secret value would change that digest, so the original claim cannot be silently rewritten. Because $\mathit{salt}$ is sampled uniformly at random and kept private, the commitment is also hiding: the digest alone reveals no information about $\mathit{candidateId}$.

Poseidon is used for this commitment step because it maps efficiently to arithmetic constraints in zk circuits \cite{poseidon2021}. That choice matters in practice: the same hash must be checked inside proof generation, so an expensive circuit hash directly increases proving overhead.

Auditability is then lifted from single commitments to the full ballot set through a Merkle tree \cite{merkle1988}. Each leaf stores one commitment, while the root commits to the entire ordered list. Inclusion proofs grow logarithmically with the number of leaves, which keeps verification cost manageable even as participation scales.

This gives two independent checks. A voter can verify that their own commitment is included under the published root without exposing additional identity information in the verification step. Separately, an external auditor can recompute the root from published entries under a deterministic ordering rule and compare it against both the announced root and the on-chain root value.

\section{Chapter Summary}

This chapter surveyed the landscape of electronic voting systems and developed the cryptographic foundations the protocol relies on. The survey established that centralised systems lack independently verifiable integrity guarantees, and that existing blockchain voting proposals address transparency but leave the privacy--verifiability integration incomplete. This dissertation targets that gap directly.

The cryptographic foundations developed in the second part---ElGamal homomorphic encryption on BabyJubJub, Groth16 zero-knowledge proofs, and Merkle-based commitment auditability---together enable a pipeline in which ballot validity and tally correctness are publicly verifiable without exposing any ballot content. Chapter~3 translates these foundations into a formal protocol model, specifying the system roles, threat assumptions, and end-to-end workflow that the implementation in Chapter~4 realises.

\endgroup
